\documentclass{ximera}
%% handout
%% nohints
%% space
%% newpage
%% numbers



\prerequisites{none}


\title{What is a Function?}

\begin{document}
\begin{abstract}
We define functions and see some examples.
\end{abstract}
\maketitle

We assume that students have encountered functions and variables. We recommend the website \link{http://www.statisticslectures.com/topics/algebraone/} as a source for students who need a more detailed description than we provide in our course. 

If two variables, $x$ and $y$, are related to each other, we say that $y$ is a \emph{function} of $x$ if each $x$ value is paired with only one $y$-value. In other words, each input value, $x$, results in exactly one output value, $y$.

For example, if $t$ is the time of day, and $s$ is the speed of your car, then is $s$ a function of $t$? At any time of day there is only one speed your car is traveling, so $s$ is a function of $t$. In other words $s$ depends on $t$.

Is $t$ a function of $s$? No. There will almost certainly be some some speed that your car reached at two different times. Thus $t$ is not a function of $s$. If you knew all the data, the speed could be determined by knowing the time, but the time could not be determined just by knowing the speed.

Another example can be found at a gas pump. Let $x$ be the money you spend on gasoline and $y$ be the number of gallons of gasoline you purchase. Then $y$ is a function of $x$ because the amount you spend determines the exact number of gallons you purchase. Also, $x$ is a function of $y$ because the amount you spend can be determined by knowing how many gallons are purchased.

The video below gives yet another example of two related variables.
\youtube{https://www.youtube.com/watch?v=D0N-DsckwtA}

\begin{question}
Select each option below where the first item is a function of the second item.

    \begin{multipleChoice}
      \choice[correct]{---the amount on a paycheck and the number of hours worked at an hourly wage job}
      \choice{---a man's height and his father's height}
      \choice{---the number of calories in a meal and the amount you pay for the meal}
      \choice[correct]{---the area of a square and the length of its sides}
      \choice[correct]{---the length of the sides of a square and its area}
      \choice{---the area of a rectangle and its perimeter}
    \end{multipleChoice}
    \begin{hint}
    If the hourly wage is set, then we should only expect one possible payout.
    \end{hint}
    \begin{hint}
    Is it possible for two fathers of the same height to have sons of different heights?
    \end{hint}
    \begin{hint}
    Is it possible that two meals that share the same price might have a different number of calories?
    \end{hint}
    \begin{hint}
    Does the formula for the area of a square depend on the side length?
    \end{hint}
    \begin{hint}
    The length of the side of a square is always the square root of the area.
    \end{hint}
    \begin{hint}
    A $3\times 5$ rectangle has the same perimeter as a $2\times 6$ rectangle. Do they have the same area?
    \end{hint}
\end{question}




\end{document}