\documentclass{ximera}
%% handout
%% nohints
%% space
%% newpage
%% numbers

\input{preamble.tex} %% we can turn off input when making a master document

\prerequisites{none}
\outcomes{ximeraLatex}

\title{Modeling and Predictions}

\begin{document}

\begin{abstract}
We model data as functions and use our models to make predictions.
\end{abstract}
\maketitle

\subsection*{Basic learning objectives}

These are the tasks you should be able to perform with reasonable fluency \textbf{when you arrive at our next class meeting}. Important new vocabulary words are indicated \emph{in italics}. 

\begin{itemize}
	\item Plot points and recognize which type of curve fits the data.
	\item Plot a curve to fit data.
\end{itemize}

\subsection*{Advanced learning objectives}

In addition to mastering the basic objectives, here are the tasks you should be able to perform \textbf{after class, with practice}: 

\begin{itemize}
	\item Use a curve fitted to a table of values to extrapolate values of a function.
    \item Be able to recognize whether a model is a good fit.
    \item Use \link{http:\\wolframalpha.com} and \link{http:\\desmos.com} to find curves that fit data.
\end{itemize}

\noindent\hrulefill

In this section we create a model by fitting a curve to data. Then we will use our model to make predictions.

Suppose we are given the function values 
\begin{center} 
    \begin{tabular}{llllllll}
    $x$  & 0 & 1 & 2 & 3 & 4 & 5 & 6 \\ \hline
    $f(x)$ & 0 & 1 & 0 & 1 & 0 & 1 & 0 \\
    \end{tabular}
\end{center}
Open up \link{desmos.com} and create a table containing the data points. A curve that that includes those points must go up and down fairly regularly, so we might try to model the curve with a wave function.

Our first guess is $f(x)=\cos(x)$ which we plot in Desmos. It doesn't work quite right, so we change the function to be $f(x)=a\cos(bx+c)+d$ and create sliders for $a$, $b$, $c$, and $d$. (Follow along by creating this function in Desmos yourself. Otherwise it won't mean anything to you.)


Our function is twice as tall as it needs to be so we set $a=.5$. The graph needs to be shifted vertically, so that its center line is at $.5$ so we set $d=.5$. Our function is too wide so we set $b=2$ or $b=3$ to make it narrower. In fact if we play with it for a minute we might decide $b=\pi$ is a good value. Then we shift things horizontally, and find that $c=\pi$ is a good choice.

Thus our model is $f(x)=.5\cos(\pi x+\pi)+.5$. Perhaps another student would have come up with $f(x)=-.5\cos(\pi x)+.5$, which happens to be the same curve. 

Is our model a good one? Probably. However, there are many other curves that fit the data.

\begin{question}
Which functions below do not fit the data for the function $f(x)$ that we have been discussing? (Please don't just guess. You won't learn anything that way.)
  \begin{solution}
    \begin{multiple-choice}
      \choice{$y=\left(\sin(\frac{3\pi}{2}(x))\right)^2$}
      \choice{$y=\big|\left|\mathrm{mod} \left(x-1,2\right)\right|-1\big|$}
      \choice{$y=\left|\sin \left(\frac{\pi x}{2}\right)\right|$}
      \choice{$y=\sin \left(2\pi x\right)-.5\cos \left(\pi x\right)+.5$}
      \choice[correct]{$y=\frac{\sin \left(\pi x\right)}{\cos \left(\pi \space x\right)+2}$}
      \choice{$y=\sin \left(\pi \left(x-2.5y\right)\right)$}
    \end{multiple-choice}
    \begin{hint}
    Enter each equation into Desmos. Be careful. 
    \end{hint}
    \begin{hint}
    You can use \verb|abs(x)| in place of \verb+|x|+ if you are struggling with the absolute value expressions. 
    \end{hint}
  \end{solution}
\end{question}

\begin{question}
Plot the data in the table
\begin{center}
\begin{tabular}{llllllll}
    $x$  & 0 & 1 & 2 & 3  & 4   & 5    & 6   \\ \hline
    $f(x)$ & 3 & 7 & 3 & -9 & -29 & -571 & -93 \\
\end{tabular}
\end{center} 
What is the parent function for the function $f(x)$?
\begin{solution}
\begin{multiple-choice}
      \choice[correct]{$y=x^2$}
      \choice{$y=x$}
      \choice{$y=x^3$}
      \choice{$y=\sqrt{x}$}
      \choice{$y=\sin(x)$}
      \choice{$y=2^x$}
      \choice{$y=\log(x)$}
\end{multiple-choice}
The formula of the function $f(x)$ is $f(x)=$ \answer{$-4x^2+8x+3$}.
\end{solution}

\end{question}




\begin{question}
Go to \link{https://student.desmos.com} and enter the code \verb|vrt3|. Use your real name, and complete the activity. We will discuss it in class.
\begin{solution}
The number of pennies in the big circle was \answer{663}.
\end{solution}
\end{question}



\end{document}