\documentclass{ximera}
%% handout
%% nohints
%% space
%% newpage
%% numbers



\title{What is probability?}

\begin{document}
\begin{abstract}
We define the concept of proability.
\end{abstract}
\maketitle

The basic idea of probability is that it measures the chance that something will happen. You may be familiar with the joke that everything is $50$-$50$ because it either happens or it doesn't! Embedded in this joke is an application of the basic formula for probability (and an example of a common error that is made).

An \emph{experiment} in probability theory is some repeatable procedure that will result in a randomized outcome. Examples of experiments include flipping a coin, rolling dice, drawing from a deck of cards, or picking balls out of an urn (any type of container will do, but mathematicians usually use urns).

If we roll a standard 6-sided die, there are six possible outcomes, which we will denote with the following notation:
\[ \{ 1, 2, 3, 4, 5, 6 \} \]

The list of all possible outcomes is known as the \emph{sample space} of an experiment. Notice that we have used curly brackets around the list and that the outcomes are separated by a comma. (This is known as a set.) An \emph{event} is some collection of outcomes that are of interest. For example, the event ``Rolling an even number'' would be the collection of events $\{ 2, 4, 6\}$ and the event ``Rolling a $1$'' would be the collection $\{ 1 \}$.

If the die has not been tampered with, then all of the outcomes are \emph{equiprobable}, meaning that the chance of each number coming up is the same. In this case, there is a simple formula that we can use to determine the probability:
\[ P(X) = \dfrac{ \text{$\#$ of successes }}{ \text{$\#$ of possibilities } } \]

In this formula, ``$P(X)$'' is read ``The probability of event $X$.'' The symbol $X$ represents a description of a type of event, so that $X$ could be phrases like ``rolling a $1$'' or ``rolling an even number.''

A \emph{success} simply means that the outcome of the experiment matches the described event, and the \# of possibilities is simply a count of all of the possible outcomes of the experiment.

Consider the following examples for the dice rolling experiment above:
\begin{itemize}
  \item If $X = $ ``rolling a $1$'' then $P(X) = \dfrac{1}{6} \approx 16.7\%$ because there is only one way to roll a 1 and six possible outcomes when rolling a die.
  \item If $X = $ ``rolling an even number'' then $P(X) = \dfrac{3}{6} = \dfrac{1}{2} = 50\%$ because there are three ways to roll an even number and six possible outcomes when rolling a die.
\end{itemize}

Sometimes we convert the fraction to a decimal or a percent, but there are many situations in which we leave it as a fraction. You will need to be able to work with all of these situations.

Now let's consider an urn that contains 999 red balls and 1 black ball. We may take the set $\{ \text{red}, \text{black} \}$ as our sample space for this experiment. 

\begin{question}
Find $P($``Choosing Black'').

    \begin{multipleChoice}
      \choice{$1/2$}
      \choice[correct]{$1/1000$}
      \end{multipleChoice}
    \begin{hint}
    The two possible outcomes are not equally likely, so $P(\text{``Choosing Black''})\ne\frac{\#\text{ of Black}}{\#\text{ of Outcomes}}$.
    \end{hint}

\end{question}

The formula $P(\text{``Choosing Black''})=\frac{\#\text{ of Black}}{\#\text{ of Outcomes}}$ only works when the outcomes are equally likely. 
Intuitively, we know that the chance of drawing a red ball is much larger than the chance of drawing a black ball. 

One way to get outcomes that are equally likely would be to label the red balls $R_1,R_2,\ldots, R_{999}$ and the black ball $B$. Then a natural sample space is $\{R_1,R_2,\ldots,R_{999},B\}$. Now each event is equally likely, so we can use the formula 
\[
P(\text{``Choosing Black''})=\frac{\#\text{ of Black}}{\#\text{ of Balls}}=\frac{1}{1000}.
\]

And this illustrates why not every event is $50$-$50$.

%%% VIDEO: Two sample problems

\begin{question}
Suppose you flip a coin. What are possible representations of the sample space of the experiment? (There are multiple correct answers.)

    \begin{multipleChoice}
      \choice{$50\%$ heads, $50\%$ tails }
      \choice[correct]{$\{ \text{heads}, \text{tails} \}$}
      \choice[correct]{$\{ \text{T}, \text{H} \}$ }
      \choice{$\{ \text{$50\%$ heads}, \text{$50\%$ tails} \}$ }
      \choice{$P(\text{tails}) = \dfrac{1}{2}, P(\text{heads}) = \dfrac{1}{2}$}
      \choice{Heads, Tails}
      \end{multipleChoice}
    \begin{hint}
    The sample space of an experiment is a list of all the possible outcomes.
    \end{hint}
    \begin{hint}
    The sample space does not include information about the probabilities.
    \end{hint}
    \begin{hint}
    Sample spaces are sets, so they require the use of set notation.
    \end{hint}

\end{question}

\begin{question}
Suppose you roll a standard 6-sided die. What is the probability of rolling a number greater than 4? (There are multiple correct answers.)

    \begin{multipleChoice}
      \choice{$\dfrac{2}{3}$}
      \choice[correct]{$\dfrac{1}{3}$}
      \choice{$\dfrac{1}{2}$}
      \end{multipleChoice}
    \begin{hint}
    $P(X) = \dfrac{ \#\text{ of successes }}{ \#\text{ of possibilities } }$
    \end{hint}
\end{question}

\begin{question}
Suppose you have an urn containing 30 red balls and 20 black balls. What is the probability of drawing a red ball?

    \begin{multipleChoice}
      \choice{$50\%$ }
      \choice{$30\%$ }
      \choice{$20\%$ }
      \choice{$75\%$ }
      \choice[correct]{$60\%$ }
      \end{multipleChoice}
    \begin{hint}
    $P(X) = \dfrac{ \#\text{ of successes }}{ \#\text{ of possibilities } }$
    \end{hint}
    \begin{hint}
    Did you try converting the fraction to a percent?
    \end{hint}

\end{question}



\end{document}
