\documentclass{ximera}
%% handout
%% nohints
%% space
%% newpage
%% numbers

\input{../preamble.tex} %% we can turn off input when making a master document




\title{Percentages}

\begin{document}
\begin{abstract}
We review the concept of percentages.
\end{abstract}
\maketitle

The word \emph{percent} means ``for each hundred.'' For example, saying Joseph earns $60$\% as much as Elaine is saying he earns \$$60$ for each \$$100$ she earns.

To calculate a percentage, the relationship between the amount that constitutes the whole, called the \emph{reference amount}, and the amount under consideration must be found. This is done via the following formula:

\begin{equation}\label{basicpercentage}
\text{Percentage}=\frac{\text{Part}}{\text{Whole}}\times 100
\end{equation}


As an example,
\[
\frac{15}{25}\times 100=60
\]
means that $15$ is $60\%$ of $25$.

The relationship in the equation $\text{Percentage}=\frac{\text{Part}}{\text{Whole}}\times 100$ can be used to find any of the three pieces---the percentage, the part, the whole---if the other two are known by using algebra to solve for the missing piece.

As an example, to find $70\%$ of $220$
\begin{align*}
70&=\frac{\text{Part}}{220}\times 100\\
70\div 100 &=\frac{\text{Part}}{220}\\
0.70\times 220 &=\text{Part}\\
154&=\text{Part}
\end{align*}
and so $70\%$ of $220$ is $154$.

\begin{question}
The label on a $12$ oz.\ can of Coca-Cola states that it contains $39$ grams of carbohydrates, accounting for $13$\% of the recommended daily value. How many grams of carbohydrates are recommended daily?

\begin{hint}
$\displaystyle 13=\frac{39}{\text{Whole}}\times 100$
\end{hint}
$\answer{300}$ grams

\end{question}


\begin{question}
For fiscal years $2013$--$2015$, the Nevada legislature appropriated $9.0\%$ of its $\$20.1$ billion budget to infrastructure. How many billions of dollars were appropriated for infrastructure? 	

\begin{hint}
\[9=\frac{\text{Part}}{\$20.1\text{ billion}}\times 100\]
\end{hint}
\begin{hint}
If you use $\$20,100,000,000$ in your calculations, your answer will be in dollars. If you use $\$20.1$ in your calculations, your answer will be in billions of dollars.
\end{hint}
\begin{hint}
Don't round your answer.
\end{hint}
\$ $\answer{1.809}$ billion

\end{question}

\begin{question}
If my taxi fare is $\$25.50$ and I pay the driver $\$30$, what percent tip am I giving?

    \begin{multipleChoice}
    	\choice{$85\%$}
        \choice{$20\%$}
        \choice[correct]{$17.6\%$}
        \choice{$15\%$}
        \choice{$1.2\%$}
    \end{multipleChoice}
    \begin{hint}
    What is the \emph{amount} of the tip? This is your \emph{part}.
    \end{hint}
    \begin{hint}
   What is the amount of the basic fare? This is your \emph{whole}.
    \end{hint}
  	\begin{hint}
    \[\text{Percentage}=\frac{\text{Part}}{\text{Whole}}\times 100\]
    \end{hint}

\end{question}




\end{document}
